\Title

### 2022-09-27 manuscript/manuscript.tex
Updated Manuscript
Update manuscript.tex

\documentclass[twocolumn]{article}

\newcommand{\Title}{\textbf{Caldera-like features located over the Panikkar Seamount and adjacent regions in the Laxmi Basin, Eastern Arabian Sea}}
\newcommand{\Journal}{Journal of Earth System Science}
\newcommand{\Author}{J. John Savio, V. Yatheesh, et al.}

\newcommand{\AuthorAffil}{
	{\large
		J. John Savio$^{1,2}$,
		V. Yatheesh$^{1,*}$,
		R. Sheldon$^{1}$,
		M. Kallathian$^{1}$,
		Pratima Kessarkar$^{1}$},\\
	{\large
		Kuldeep Kumar$^{1}$,
		P. Mahesh$^{1}$,
		Subhashree Mishra$^{1}$,
		and Swati Verma$^{1}$
	}
	\\[0.4cm]
	{\small $^1$ CSIR - National Institute of Oceanography, Dona Paula, Goa -  403 004, India}
	\\
	{\small $^2$ School of Earth, Ocean and Atmospheric Sciences, Goa University, Taleigao Plateau, Goa -  403 206, India}
	\\
	{\small $^*$ Corresponding author e-mail: yatheesh@nio.org}
}
\newcommand{\DOI}{\href{https://doi.org/10.1007/s12040-021-01797-7}{10.1007/s12040-021-01797-7}}
\newcommand*{\doi}[1]{doi:\href{http://dx.doi.org/#1}{#1}} %% To give hyperlink to DOI's

\usepackage[left=0.7in,right=0.7in,top=0.9in,bottom=0.8in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{ragged2e}
\usepackage{graphicx}
\usepackage[round]{natbib}
\hbadness=99999
%%\usepackage{fixltx2e}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
	allcolors=blue,
	pdftitle={\Title},
	pdfauthor={\Author},
}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{caption}
\fancyhf{}
\lhead{
	\fontsize{9pt}{12pt}\selectfont
	\Author{}, \DOI{}
}
\rhead{\fontsize{9pt}{12pt}\selectfont \thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}
\title{\Title}
\author{\AuthorAffil}
\date{
\normalsize
MS received - 12 May 2021; revised - 27 October 2021; accepted - 05 November 2021
\\[0.4cm] 
\justifying{
This is a pre-copyedited, author-produced version of an article accepted for publication in \textit{\Journal{}} following peer review. The version of record "\textit{John Savio, J., Yatheesh, V., Sheldon, R., Kallathian, M., Kessarkar, P., Kumar, K., Mahesh, P., Mishra, S., \&  Verma, S., 2022. \Title{}, \Journal{},}\ " is available online at: \DOI{}
}}
\maketitle

@@

The axial part of the Laxmi Basin is characterized by the presence of three isolated seamounts, referred to as the Raman Seamount, Panikkar Seamount and Wadia Guyot \citep{Bhattacharya1994b}, from the north to south (Figures \ref{fig:studyarea} \& \ref{fig:studyarea2}). These seamounts, together forming an axial seamount chain in the Laxmi Basin \citep{Bhattacharya1994b}, situates on an ~ 360 km long, NNW-SSE striking ridge buried under the sediments (Figures \ref{fig:studyarea} \& \ref{fig:studyarea2}), referred to as the Panikkar Ridge \citep{Gopala_Rao1992}. The supporters of continental origin hypothesis for the formation of the Laxmi Basin interpreted the Panikkar Ridge to\begin{figure*}[!htb]

@@

\textsl{	Topographic map of the northwestern continental margin of India and the adjoining deep oceans basins as depicted by the GEBCO 2021 global terrain model for ocean and land \citep{GEBCO_Compilation_Group2021}, showing the locations of the Panikkar Ridge (shown as thick white line) and the axial seamount chain in the Laxmi Basin. Thin black lines are selected (1000 m., 2000 m., 3000 m., and 4000 m.) bathymetry contours from GEBCO 2021 data set. The box shown with thick dashed black lines represents the boundary of the tectonic map shown in Figure \ref{fig:studyarea2}. LXR: Laxmi Ridge; LXB: Laxmi Basin; ABB: Arabian Basin; PKR: Panikkar Ridge; LCP: Laccadive Plateau. R: Raman Seamount; P: Panikkar Seamount; W: Wadia Guyot}

@@

\end{figure*} represent a structural feature of less altered stretched continental crust \citep{Krishna2006}, or an isolated continental block \citep{Geoffroy2020}. \cite{Krishna2006} further considered that the Raman Seamount, Panikkar Seamount and Wadia Guyot represent the isolated highs of the Panikkar Ridge elevated above the seafloor, implying a continental origin for these seamounts. In contrary to this, the supporters of oceanic crust hypothesis for the formation of the Laxmi Basin inferred that the Panikkar Ridge represents the extinct spreading centre in the Laxmi Basin \citep{Bhattacharya1994a,Bhattacharya1994b}, where two-limbed seafloor spreading initiated at around 67.6 Ma and ceased at ~56.4 Ma, as interpreted from the identified magnetic anomaly sequence C30n-C25r-C30n \citep{Bhattacharya2015,Yatheesh2020}. Further, those studies attributed the genesis of the Raman-Panikkar-Wadia seamount chain located over the Panikkar Ridge to an anomalous episode of volcanism resulting from the interaction of the R\'eunion hotspot with the spreading centre in the Laxmi Basin. These seamounts are considered to have been emplaced at around 62.5 Ma \citep{Bhattacharya2015,Yatheesh2020}, coeval with the emplacement of Ghatkopar-Powai tholeiites on the Indian mainland \citep{Pande2017}. However, any direct evidences from this seamount chain to represent their genesis in terms of volcanic eruptions is still awaited. Towards this, we carried out a multibeam bathymetry survey over a selected sector of this seamount chain to

@@

understand the detailed seafloor morphology of this anomalous feature. In this paper, we present the preliminary morphotectonic information derived from the analysis of this fresh set of recently acquired high-resolution multibeam bathymetry data.

@@

We present high-resolution multibeam bathymetry data as a colour-coded two-dimensional (Figures \ref{panikkar}a\&b) and three-dimensional (Figures \ref{calderas}a-c) images over a segment of the Laxmi Basin Seamount chain, i.e., over the Panikkar Seamount and the adjoining regions. The multibeam bathymetric map (Figure \ref{panikkar}a) clearly depicts the presence of three bathymetric highs, together forming an arcuate shape. Among these, the two larger features with major peaks sitting on a single base defined by 3600 m. isobath, represent the Panikkar Seamount discovered by \cite{Bhattacharya1994b}. The third bathymetric high identified in the present study also appears to form a part of this anomalous feature, and therefore, Panikkar Seamount can be considered to consists of three segments. For the ease of reference, in this paper, we refer these segments as Segment-1 (S1), Segment-2 (S2) and Segment-3 (S3) of the Panikkar Seamount, from northwest to southeast, respectively (Figure \ref{panikkar}a). Among these, the Segment-1 (S1) represents the largest feature, whose western side is characterized by a sharp boundary defined by the 3650 m. and 3600 m. isobaths. In the eastern side of this segment, north of 16°09’N, the bathymetric high shows similar sharp drop in the depth as in the case of western limit, however, towards south, this feature shows a relatively gentle topography. Similarly, the Segment-2 (S2) is associated with a sharp boundary in the southwestern side and a relatively gentle topography on its eastern side. However, the bathymetric map of the regions adjacent to these two segments, Segment-1 and Segment-2, clearly depicts the presence of isolated, characteristic semi-circular, scarp faces / embayments with a moderately steep slope that are dipping to a nearly flat seafloor region away from the crest of the seamount, in the southwestern as well as eastern sides of these anomalous bathymetric features (Figure \ref{panikkar}a). Similarly, the Segment-3, which is separated from Segment-2 by a relatively deeper seafloor, is also characterized by the presence of two smaller semi-circular bathymetric features (Figure \ref{panikkar}a). Morphologically, such rim-like features enclosing a region of nearly flat seafloor associated with all the three segments of the Panikkar Ridge are very interesting, since this geometry closely resembles with the shape of volcanic caldera structures associated with the submarine volcanic events. Such volcanic calderas generally form due to the collapse or subsidence into the top of a magma chamber \citep{Cole2005}. \cite{Bijesh2021} has recently reported the presence of an extinct giant volcanic caldera in the Arabian Sea, southwest of the Laccadive Plateau, based on the presence of a nearly elliptical bathymetry high complex, known as the Sagar Kanya Bathymetric High Complex, enclosing a nearly flat seafloor region.
\begin{minipage}{\linewidth}% to keep image and caption on one page

@@

@@

It may be noted that the inferred caldera-like features are concentrated south of 16° 08’N latitude on the Panikkar Seamount, while the northern part of this seamount is devoid of such prominent features. In the north, the Segment S-1 contains multiple peaks on a single base and the least depth over the largest of these peaks (near 16°13'N) is about 2618 m., which is much higher from the surrounding seafloor than the other two segments, suggesting multiple phases of volcanism over the Segment S-1 of the Panikkar Seamount. As in the case of region south of 16° 08’N latitude, volcanic caldera-like features might have been formed in this norther region also, however, the later phases of volcanism occurred in the northernmost part of the Panikkar Seamount might have obliterated the morphological features representing caldera-like feature formed as a result of the early phases of volcanism.   
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass[twocolumn]{article}
\newcommand{\Title}{\textbf{Caldera-like features located over the Panikkar Seamount and adjacent regions in the Laxmi Basin, Eastern Arabian Sea}}
\newcommand{\Journal}{Journal of Earth System Science}
\newcommand{\Author}{J. John Savio, V. Yatheesh et al. (2022)}

\newcommand{\AuthorAffil}{
	{\large
		J. John Savio$^{1,2}$,
		V. Yatheesh$^{1,*}$,
		R. Sheldon$^{1}$,
		M. Kallathian$^{1}$,
		Pratima Kessarkar$^{1}$},\\
	{\large
		Kuldeep Kumar$^{1}$,
		P. Mahesh$^{1}$,
		Subhashree Mishra$^{1}$,
		and Swati Verma$^{1}$
	}
	\\[0.4cm]
	{\small $^1$ CSIR - National Institute of Oceanography, Dona Paula, Goa -  403 004, India}
	\\
	{\small $^2$ School of Earth, Ocean and Atmospheric Sciences, Goa University, Taleigao Plateau, Goa -  403 206, India}
	\\
	{\small $^*$ Corresponding author e-mail:\hspace{2mm} yatheesh@nio.org \hspace{2mm} yatheeshv@gmail.com}
}
\newcommand{\DOI}{\href{https://doi.org/10.1007/s12040-021-01797-7}{10.1007/s12040-021-01797-7}}
\newcommand*{\doi}[1]{doi:\href{http://dx.doi.org/#1}{#1}} %% To give hyperlink to DOI's

\usepackage[left=0.7in,right=0.7in,top=0.9in,bottom=0.8in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{ragged2e}
\usepackage{graphicx}
\usepackage[round]{natbib}
\hbadness=99999
%%\usepackage{fixltx2e}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
	allcolors=blue,
	pdftitle={\Title},
	pdfauthor={\Author},
	pdfpagemode=UseNone
}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{caption}
\fancyhf{}
\lhead{
	\fontsize{9pt}{12pt}\selectfont
	\Author{}, \DOI{}
}
\rhead{\fontsize{9pt}{12pt}\selectfont \thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}
\title{\Title}
\author{\AuthorAffil}
\date{
\normalsize
Manuscript received - 12 May 2021; Manuscript revised - 27 October 2021; \\ Accepted - 05 November 2021; Final version published - 02 February 2022
\\[0.4cm] 
\justifying{
This is a pre-copyedited, author-produced version of an article accepted for publication in \textit{\Journal{}} following peer review. The version of record "\textit{John Savio, J., Yatheesh, V., Sheldon, R., Kallathian, M., Kessarkar, P., Kumar, K., Mahesh, P., Mishra, S., \&  Verma, S., 2022. \Title{}, \Journal{},}\ " is available online at: \DOI{}
}}
\maketitle

@@

@@

		\textsl{	Topographic map of the northwestern continental margin of India and the adjoining deep oceans basins as depicted by the GEBCO 2021 global terrain model for ocean and land \citep{GEBCO_Compilation_Group2021}, showing the locations of the Panikkar Ridge (shown as thick white line) and the axial seamount chain in the Laxmi Basin. Thin black lines are selected (1000 m., 2000 m., 3000 m., and 4000 m.) bathymetry contours from GEBCO 2021 data set. The box shown with thick dashed black lines represents the boundary of the tectonic map shown in Figure \ref{fig:studyarea2}. LXR: Laxmi Ridge; LXB: Laxmi Basin; ABB: Arabian Basin; PKR: Panikkar Ridge; LCP: Laccadive Plateau. R: Raman Seamount; P: Panikkar Seamount; W: Wadia Guyot.}

@@

The axial part of the Laxmi Basin is characterized by the presence of three isolated seamounts, referred to as the Raman Seamount, Panikkar Seamount and Wadia Guyot \citep{Bhattacharya1994b}, from the north to south (Figures \ref{fig:studyarea} \& \ref{fig:studyarea2}). These seamounts, together forming an axial seamount chain in the Laxmi Basin \citep{Bhattacharya1994b}, situates on an {$\sim$} 360 km long, NNW-SSE striking ridge buried under the sediments (Figures \ref{fig:studyarea} \& \ref{fig:studyarea2}), referred to as the Panikkar Ridge \citep{Gopala_Rao1992}. The supporters of continental origin hypothesis for the formation of the Laxmi Basin interpreted the Panikkar Ridge to represent a structural feature of less altered stretched continental crust \citep{Krishna2006}, or an isolated continental block \citep{Geoffroy2020}. \cite{Krishna2006} further considered that the Raman Seamount, Panikkar Seamount and Wadia Guyot represent the isolated highs of the Panikkar Ridge elevated above the seafloor, implying a continental origin for these seamounts.

@@

In contrary to this, the supporters of oceanic crust hypothesis for the formation of the Laxmi Basin inferred that the Panikkar Ridge represents the extinct spreading centre in the Laxmi Basin \citep{Bhattacharya1994a,Bhattacharya1994b}, where two-limbed seafloor spreading initiated at around 67.6 Ma and ceased at ~56.4 Ma, as interpreted from the identified magnetic anomaly sequence C30n-C25r-C30n \citep{Bhattacharya2015,Yatheesh2020}. Further, those studies attributed the genesis of the Raman-Panikkar-Wadia seamount chain located over the Panikkar Ridge to an anomalous episode of volcanism resulting from the interaction of the R\'eunion hotspot with the spreading centre in the Laxmi Basin. These seamounts are considered to have been emplaced at around 62.5 Ma \citep{Bhattacharya2015,Yatheesh2020}, coeval with the emplacement of Ghatkopar-Powai tholeiites on the Indian mainland \citep{Pande2017}. However, any direct evidences from this seamount chain to represent their genesis in terms of volcanic eruptions is still awaited. Towards this, we carried out a multibeam bathymetry survey over a selected sector of this seamount chain tounderstand the detailed seafloor morphology of this anomalous feature. In this paper, we present the preliminary morphotectonic information derived from the analysis of this fresh set of recently acquired high-resolution multibeam bathymetry data.

@@

We present high-resolution multibeam bathymetry data as a colour-coded two-dimensional (Figures \ref{panikkar}a\&b) and three-dimensional (Figures \ref{calderas}a-c) images over a segment of the Laxmi Basin Seamount chain, i.e., over the Panikkar Seamount and the adjoining regions. The multibeam bathymetric map (Figure \ref{panikkar}a) clearly depicts the presence of three bathymetric highs, together forming an arcuate shape. Among these, the two larger features with major peaks sitting on a single base defined by 3600 m. isobath, represent the Panikkar Seamount discovered by \cite{Bhattacharya1994b}. The third bathymetric high identified in the present study also appears to form a part of this anomalous feature, and therefore, Panikkar Seamount can be considered to consists of three segments. For the ease of reference, in this paper, we refer these segments as Segment-1 (S1), Segment-2 (S2) and Segment-3 (S3) of the Panikkar Seamount, from northwest to southeast, respectively (Figure \ref{panikkar}a). Among these, the Segment-1 (S1) represents the largest feature, whose western side is characterized by a sharp boundary defined by the 3650 m. and 3600 m. isobaths. In the eastern side of this segment, north of 16°09’N, the bathymetric high shows similar sharp drop in the depth as in the case of western limit, however, towards south, this feature shows a relatively gentle topography. Similarly, the Segment-2 (S2) is associated with a sharp boundary in the southwestern side and a relatively gentle topography on its eastern side. However, the bathymetric map of the regions adjacent to these two segments, Segment-1 and Segment-2, clearly depicts the presence of isolated, characteristic semi-circular, scarp faces / embayments with a moderately steep slope that are dipping to a nearly flat seafloor region away from the crest of the seamount, in the southwestern as well as eastern sides of these anomalous bathymetric features (Figure \ref{panikkar}a). Similarly, the Segment-3, which is separated from Segment-2 by a relatively deeper seafloor, is also characterized by the presence of two smaller semi-circular bathymetric features (Figure \ref{panikkar}a). Morphologically, such rim-like features enclosing a region of nearly flat seafloor associated with all the three segments of the Panikkar Ridge are very interesting, since this geometry closely resembles with the shape of volcanic caldera structures associated with the submarine volcanic events. Such volcanic calderas generally form due to the collapse or subsidence into the \begin{minipage}{\linewidth}% to keep image and caption on one page

@@

top of a magma chamber \citep{Cole2005}. \cite{Bijesh2021} has recently reported the presence of an extinct giant volcanic caldera in the Arabian Sea, southwest of the Laccadive Plateau, based on the presence of a nearly elliptical bathymetry high complex, known as the Sagar Kanya Bathymetric High Complex, enclosing a nearly flat seafloor region.

@@

It may be noted that the inferred caldera-like features are concentrated south of 16° 08’N latitude on the Panikkar Seamount, while the northern part of this seamount is devoid of such prominent features. In the north, the Segment S1 contains multiple peaks on a single base and the least depth over the largest of these peaks (near 16°13'N) is about 2618 m., which is much higher from the surrounding seafloor than the other two segments, suggesting multiple phases of volcanism over the Segment S1 of the Panikkar Seamount. As in the case of region south of 16° 08’N latitude, volcanic caldera-like features might have been formed in this norther region also, however, the later phases of volcanism occurred in the northernmost part of the Panikkar Seamount might have obliterated the morphological features formed as a result of the early phases of volcanism.   
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
